---
title: "Structural modeling"
author: "Guilhem Huau"
date: "`r format(Sys.time(), '%d/%m/%Y %H:%M:%S')`"
colorlinks: true
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: TRUE
  pdf_document:
    toc: true
    toc_depth: 4
    number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      include = TRUE,
                      warning = TRUE,
                      message = TRUE,
                      
                      cache = FALSE, # save chunks output 
                      
                      results = "hold", # display chunk output in one block
                      
                      tidy = FALSE, # better code organisation
                      tidy.opts = list(width.cutoff = 80),
                      strip.white = TRUE, #remove the white lines in the beginning or end of a source chunk
                      
                      fig.align = "center",
                      fig.width = 9,
                      fig.height = 5)

knitr::opts_knit$set(progress = TRUE,
                     verbose = FALSE,
                     width = 80)

```

```{css echo=FALSE, eval = TRUE}
/* add some CSS styling here */
```

**Packages / R-options:**
```{r pkgs_options, message=FALSE, warning=FALSE}
suppressPackageStartupMessages({
  require(dplyr)
  require(tibble)
  require(tidyr)
  require(magrittr)
  require(lavaan)
  require(lavaanPlot)
  require(plotly)
  require(FactoMineR)
})

#  R options
options(stringsAsFactors = FALSE)

```

***

```{r loading_data, include=FALSE}
dat_root = read.csv("2017_Tottori_May_RootPhenotype.csv")
dat_shoot = read.csv("2017_Tottori_May_ShootPhenotype.csv")
dat_shoot$year = 2017
dat_join = inner_join(dat_shoot, dat_root, by = colnames(dat_root)[1:9])
dat_join$ID = 1:nrow(dat_join)
dat_join$block = as.factor(dat_join$block)
weightProblem = filter(dat_join,LeafWeight_g > TotalWeight_g)
dat_join = anti_join(dat_join, weightProblem)

dat_join = dat_join[,-(c(1:6,8,9))]

dat_complete = dat_join[complete.cases(dat_join),]
```

```{r model, include=FALSE}

dat_choose = dat_complete[,c(1:5,9)]
names(dat_choose) = c("block", "PlantHeight","LeafWeight","TotalWeight","TRL","NumTips")
# dat_test = filter(dat_test, block == "D")


dat_test = mutate_if(dat_choose, is.numeric, scale)

mtest = 'root =~ TRL + NumTips
        shoot =~ LeafWeight + PlantHeight + TotalWeight
        shoot ~ root
        TotalWeight ~~ LeafWeight + PlantHeight'

fit_test = cfa(model = mtest, data = dat_test, group = "block")
```

# Model




On the figure the arrow and the number on the left stand for the C(ontrol) group, while the ones on the right are the D(rought) group.

```{r model_plot, echo=FALSE}
labels = list(root = "Root", shoot = "Shoot", TRL = "Total Root Length", NumTips = "Number of Tips", LeafWeight = "Leaf Weight", PlantHeight = "Plant Height", TotalWeight = "Total Weight")
lavaanPlot(name = "Model", model = fit_test, labels = labels, coefs = TRUE)

```


***

# Fit Summary


The first part of this output give various informations about the model and how good it is. After you can read the results for each group.

```{r}
summary(fit_test, standardized=TRUE)
```

```{r echo=FALSE}
lat_data = predict(fit_test, dat_test) #to get the latent values for each individuals

dat_lat1 = lat_data$C
dat_lat1 = data.frame(block = "C", dat_lat1)

dat_lat2 = lat_data$D
dat_lat2 = data.frame(block = "D", dat_lat2)

a = bind_rows(dat_lat1, dat_lat2)

dat_ordered = arrange(dat_choose, block)
b = bind_cols(dat_ordered,a)[,-7]

```



***
# General statistics


```{r general, echo=FALSE}
summary(b)

```

***

# Group statistics

***

## Plant Height Statistics


```{r plantHeight, echo=FALSE}
tapply(b[,2], b$block, summary)
plot_ly(b, x = ~block, y= ~PlantHeight, type = "box")
```


***

## Leaf Weight Statistics


```{r LeafWeight, echo=FALSE}
tapply(b[,3], b$block, summary)
plot_ly(b, x = ~block, y= ~LeafWeight, type = "box")
```


***

## Total Weight Statistics


```{r TotalWeight, echo=FALSE}
tapply(b[,4], b$block, summary)
plot_ly(b, x = ~block, y= ~TotalWeight, type = "box")
```


***

## Total Root Length Statistics


```{r TRL, echo=FALSE, warning=FALSE}
tapply(b[,5], b$block, summary)
plot_ly(b, x = ~block, y= ~TRL, type = "box")
```


***

## Number of Root Tips Statistics


```{r NumTips, echo=FALSE}
tapply(b[,6], b$block, summary)
plot_ly(b, x = ~block, y= ~NumTips, type = "box")
```

***

## Shoot (latent variable) Statistics


```{r Shoot, echo=FALSE}
tapply(b[,7], b$block, summary)
plot_ly(b, x = ~block, y= ~shoot, type = "box")
```

***

## Root (latent variable) Statistics


```{r Root, echo=FALSE}
tapply(b[,8], b$block, summary)
plot_ly(b, x = ~block, y= ~root, type = "box")
```

***

# Root - Shoot Plot


Note: Root and shoot are standardized variables.

```{r root_shoot_plot, echo=FALSE, warning=FALSE}
fit_root_shoot_C = lm(data = filter(b, block == "C"), shoot ~ root)
fit_root_shoot_D = lm(data = filter(b, block == "D"), shoot ~ root)

x <- list(title = "Root")
y <- list(title = "Shoot")
plot_ly(b, x= ~root, y= ~shoot, color = ~block, type = "scatter", mode = "markers") %>%
  layout(title = "Shoot(Root) and linear regressions for each group", xaxis = x, yaxis = y) %>%
  add_lines(data = filter(b, block == "C"), y = fitted(fit_root_shoot_C), text = paste("shoot =", round(fit_root_shoot_C$coefficients[2], 3), "* root"), hoverinfo = "text") %>%
  add_lines(data = filter(b, block == "D"), y = fitted(fit_root_shoot_D), text = paste("shoot =", round(fit_root_shoot_D$coefficients[2], 3), "* root"), hoverinfo = "text")

print(paste("Correlation between fit and value for group C:",round(cor(b$shoot[which(b$block=="C")],fitted(fit_root_shoot_C)),3)))
print(paste("Correlation between fit and value for group D:",round(cor(b$shoot[which(b$block=="D")],fitted(fit_root_shoot_D)),3)))
```

(You can click on the legend to choose which layers you want to see)

***

# PCA Analysis

## PCA on Root variables only

When Comp1 and Comp2 (and Comp4 and Comp5) are not both in the model, no error (but still p-value = 0 -> strange)

```{r PCA}

PCA_root = PCA(dat_complete, scale.unit=TRUE, ncp=5, graph=F, quali.sup = 1, quanti.sup = c(2:4))
PCA_root_coord = PCA_root[["var"]][["coord"]]

dat_PCA_root = data.frame(as.matrix.data.frame(dat_complete[,-c(1:4)]) %*% as.matrix.data.frame(PCA_root_coord))
dat_PCA_with_block = data.frame(dat_complete[,c(1:4)], dat_PCA_root)
dat_fit_PCA = mutate_if(dat_PCA_with_block, is.numeric, scale)

names(dat_fit_PCA) = c("block","PlantHeight","LeafWeight","TotalWeight","CompR1","CompR2","CompR3","CompR4","CompR5")

mPCA = 'root =~ CompR1 + CompR2
        shoot =~ LeafWeight + PlantHeight + TotalWeight
        shoot ~ root
        TotalWeight ~~ LeafWeight'

fit_PCA = cfa(model = mPCA, data = dat_fit_PCA, group = "block")

summary(fit_PCA)
```

```{r predict_PCA, echo=FALSE}
lat_data = predict(fit_PCA, dat_fit_PCA) #to get the latent values for each individuals

dat_lat1 = lat_data$C
dat_lat1 = data.frame(block = "C", dat_lat1)

dat_lat2 = lat_data$D
dat_lat2 = data.frame(block = "D", dat_lat2)

a = bind_rows(dat_lat1, dat_lat2)

dat_ordered = arrange(dat_choose, block)
b = bind_cols(dat_ordered,a)[,-7]

```

```{r root_shoot_plot_PCA, echo=FALSE, warning=FALSE}
fit_root_shoot_C = lm(data = filter(b, block == "C"), shoot ~ root)
fit_root_shoot_D = lm(data = filter(b, block == "D"), shoot ~ root)

x <- list(title = "Root")
y <- list(title = "Shoot")
plot_ly(b, x= ~root, y= ~shoot, color = ~block, type = "scatter", mode = "markers") %>%
  layout(title = "Shoot(Root) and linear regressions for each group", xaxis = x, yaxis = y) %>%
  add_lines(data = filter(b, block == "C"), y = fitted(fit_root_shoot_C), text = paste("shoot =", round(fit_root_shoot_C$coefficients[2], 3), "* root"), hoverinfo = "text") %>%
  add_lines(data = filter(b, block == "D"), y = fitted(fit_root_shoot_D), text = paste("shoot =", round(fit_root_shoot_D$coefficients[2], 3), "* root"), hoverinfo = "text")

print(paste("Correlation between fit and value for group C:",round(cor(b$shoot[which(b$block=="C")],fitted(fit_root_shoot_C)),3)))
print(paste("Correlation between fit and value for group D:",round(cor(b$shoot[which(b$block=="D")],fitted(fit_root_shoot_D)),3)))
```

## PCA on Root and Shoot variables

When Comp1 and Comp2 (and Comp4 and Comp5) are not both in the model, no error (but still p-value = 0 -> strange)

```{r PCA2}
PCA_shoot = PCA(dat_complete, scale.unit=TRUE, ncp=5, graph=F, quali.sup = 1, quanti.sup = c(5:24))
PCA_shoot_coord = PCA_shoot[["var"]][["coord"]]
dat_PCA_shoot = data.frame(as.matrix.data.frame(dat_complete[,-c(1,5:24)]) %*% as.matrix.data.frame(PCA_shoot_coord))

dat_PCA_with_block2 = data.frame(block = dat_complete[,1], dat_PCA_shoot, dat_PCA_root)
dat_fit_PCA2 = mutate_if(dat_PCA_with_block2, is.numeric, scale)

names(dat_fit_PCA2) = c("block","CompS1","CompS2","CompS3","CompR1","CompR2","CompR3","CompR4","CompR5")

mPCA2 = 'root =~ CompR1 + CompR3
        shoot =~ CompS1 + CompS2
        shoot ~ root'

fit_PCA2 = cfa(model = mPCA2, data = dat_fit_PCA2, group = "block")

summary(fit_PCA2)
```

```{r predict_PCA2, echo=FALSE}
lat_data = predict(fit_PCA2, dat_fit_PCA2) #to get the latent values for each individuals

dat_lat1 = lat_data$C
dat_lat1 = data.frame(block = "C", dat_lat1)

dat_lat2 = lat_data$D
dat_lat2 = data.frame(block = "D", dat_lat2)

a = bind_rows(dat_lat1, dat_lat2)

dat_ordered = arrange(dat_choose, block)
b = bind_cols(dat_ordered,a)[,-7]

```

```{r root_shoot_plot_PCA2, echo=FALSE, warning=FALSE}
fit_root_shoot_C = lm(data = filter(b, block == "C"), shoot ~ root)
fit_root_shoot_D = lm(data = filter(b, block == "D"), shoot ~ root)

x <- list(title = "Root")
y <- list(title = "Shoot")
plot_ly(b, x= ~root, y= ~shoot, color = ~block, type = "scatter", mode = "markers") %>%
  layout(title = "Shoot(Root) and linear regressions for each group", xaxis = x, yaxis = y) %>%
  add_lines(data = filter(b, block == "C"), y = fitted(fit_root_shoot_C), text = paste("shoot =", round(fit_root_shoot_C$coefficients[2], 3), "* root"), hoverinfo = "text") %>%
  add_lines(data = filter(b, block == "D"), y = fitted(fit_root_shoot_D), text = paste("shoot =", round(fit_root_shoot_D$coefficients[2], 3), "* root"), hoverinfo = "text")

print(paste("Correlation between fit and value for group C:",round(cor(b$shoot[which(b$block=="C")],fitted(fit_root_shoot_C)),3)))
print(paste("Correlation between fit and value for group D:",round(cor(b$shoot[which(b$block=="D")],fitted(fit_root_shoot_D)),3)))
```


## PCA on Root and Ionomics variables

### Fitting and summary of the model

When Comp1 and Comp2 (and Comp4 and Comp5) are not both in the model, no error (but still p-value = 0 -> strange)

```{r loading_data_iono, message=FALSE, warning=FALSE}
# dat ion ICP = data only for one individual per variety
# dat_ionICP = read.csv("2017_Tottori_May_IonomeICP.csv")
dat_ionX = read.csv("2017_Tottori_May_IonomeXray.csv")

dat_shoot$year = 2017
dat_join = inner_join(dat_shoot, dat_root, by = colnames(dat_root)[1:9])
weightProblem = filter(dat_join,LeafWeight_g > TotalWeight_g)
dat_join = anti_join(dat_join, weightProblem)
dat_IandP = inner_join(dat_join, dat_ionX, by = colnames(dat_join)[1:9])
dat_IandP$block = as.factor(dat_IandP$block)
dat_IandP$ID = 1:nrow(dat_IandP)

dat_IandP = dat_IandP[,-(c(1:6,8,9))]

dat_complete2 = dat_IandP[complete.cases(dat_IandP),]


```

```{r PCA3}
PCA_iono = PCA(dat_complete2, scale.unit=TRUE, ncp=5, graph=F, quali.sup = 1, quanti.sup = c(2:24))
PCA_iono_coord = PCA_iono[["var"]][["coord"]]
dat_PCA_iono = data.frame(as.matrix.data.frame(dat_complete2[,-c(1,1:24)]) %*% as.matrix.data.frame(PCA_iono_coord))

PCA_root2 = PCA(dat_complete2, scale.unit=TRUE, ncp=5, graph=F, quali.sup = 1, quanti.sup = c(2:4,25:43))
PCA_root_coord2 = PCA_root2[["var"]][["coord"]]
dat_PCA_root2 = data.frame(as.matrix.data.frame(dat_complete2[,-c(1:4,25:43)]) %*% as.matrix.data.frame(PCA_root_coord2))

dat_PCA_with_block3 = data.frame(dat_complete2[,c(1:4)], dat_PCA_root2, dat_PCA_iono)
dat_fit_PCA3 = mutate_if(dat_PCA_with_block3, is.numeric, scale)

names(dat_fit_PCA3) = c("block","PlantHeight","LeafWeight","TotalWeight","CompR1","CompR2","CompR3","CompR4","CompR5","CompI1","CompI2","CompI3","CompI4","CompI5")

mPCA3 = 'root =~ CompR1 + CompR3
        shoot =~ LeafWeight + PlantHeight + TotalWeight
        iono =~ CompI1  + CompI3 + CompI4 + CompI5
        shoot ~ root + iono
        TotalWeight ~~ LeafWeight'

fit_PCA3 = cfa(model = mPCA3, data = dat_fit_PCA3, group = "block")

summary(fit_PCA3)
```

***

### Graph of Shoot/Root and Shoot/Ionomics

```{r predict_PCA3, echo=FALSE, warning=FALSE}
lat_data = predict(fit_PCA3, dat_fit_PCA3) #to get the latent values for each individuals

dat_lat1 = lat_data$C
dat_lat1 = data.frame(block = "C", dat_lat1)

dat_lat2 = lat_data$D
dat_lat2 = data.frame(block = "D", dat_lat2)

a = bind_rows(dat_lat1, dat_lat2)

dat_choose2 = dat_complete2[,c(1:5,9)]
dat_ordered2 = arrange(dat_choose2, block)
b = bind_cols(dat_ordered2,a)[,-7]

```

```{r root_shoot_plot_PCA3, echo=FALSE, warning=FALSE}
fit_root_shoot_C = lm(data = filter(b, block == "C"), shoot ~ root + iono)
fit_root_shoot_D = lm(data = filter(b, block == "D"), shoot ~ root + iono)

x <- list(title = "Root")
y <- list(title = "Shoot")
plot_ly(b, x= ~root, y= ~shoot, color = ~block, type = "scatter", mode = "markers") %>%
  layout(title = "Shoot(Root) and linear regressions for each group", xaxis = x, yaxis = y) %>%
  add_lines(data = filter(b, block == "C"), y = fitted(fit_root_shoot_C), text = paste("shoot =", round(fit_root_shoot_C$coefficients[2], 3), "* root"), hoverinfo = "text") %>%
  add_lines(data = filter(b, block == "D"), y = fitted(fit_root_shoot_D), text = paste("shoot =", round(fit_root_shoot_D$coefficients[2], 3), "* root"), hoverinfo = "text")

print(paste("Correlation between fit and value for group C:",round(cor(b$shoot[which(b$block=="C")],fitted(fit_root_shoot_C)),3)))
print(paste("Correlation between fit and value for group D:",round(cor(b$shoot[which(b$block=="D")],fitted(fit_root_shoot_D)),3)))
```


```{r iono_shoot_plot_PCA3, echo=FALSE, warning=FALSE}
fit_root_shoot_C = lm(data = filter(b, block == "C"), shoot ~ root + iono)
fit_root_shoot_D = lm(data = filter(b, block == "D"), shoot ~ root + iono)

x <- list(title = "Iono")
y <- list(title = "Shoot")
plot_ly(b, x= ~iono, y= ~shoot, color = ~block, type = "scatter", mode = "markers") %>%
  layout(title = "Shoot(Root) and linear regressions for each group", xaxis = x, yaxis = y) %>%
  add_lines(data = filter(b, block == "C"), y = fitted(fit_root_shoot_C), text = paste("shoot =", round(fit_root_shoot_C$coefficients[2], 3), "* iono"), hoverinfo = "text") %>%
  add_lines(data = filter(b, block == "D"), y = fitted(fit_root_shoot_D), text = paste("shoot =", round(fit_root_shoot_D$coefficients[2], 3), "* iono"), hoverinfo = "text")

print(paste("Correlation between fit and value for group C:",round(cor(b$shoot[which(b$block=="C")],fitted(fit_root_shoot_C)),3)))
print(paste("Correlation between fit and value for group D:",round(cor(b$shoot[which(b$block=="D")],fitted(fit_root_shoot_D)),3)))
```


***

### Shoot (latent variable) Statistics


```{r Shoot2, echo=FALSE}
tapply(b[,7], b$block, summary)
plot_ly(b, x = ~block, y= ~shoot, type = "box")
```

***

### Root (latent variable) Statistics


```{r Root2, echo=FALSE}
tapply(b[,8], b$block, summary)
plot_ly(b, x = ~block, y= ~root, type = "box")
```

***

# Appendix
```{r sessionInfo}
print(sessionInfo(), locale = FALSE)
```
